---
title: 'Reproducible Research: Peer Assessment 1'
author: "Luis Fernando Agottani"
date: "09/08/2019"
output:
  html_document: default
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Verify if the filename has been downloaded/unzip yet. If not, download/unzip the file with the fileURL link.

```{r Loanding Data, echo=TRUE}
filename <- "repdata_data_activity.zip"

if (!file.exists(filename)){
  fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file(fileURL, filename)
}
if (!file.exists("activity.csv")) { 
  unzip(filename)
}
```

Reading data.csv.

```{r Preprocessing the data, echo=TRUE}

Act<- read.csv("activity.csv", 
               head = TRUE, sep= ",")
```

Changing the Factor type of the column date to Date with as.Date.

```{r Cleaning the data, echo=TRUE}

Act$date<- as.Date(Act$date)
```

## What is mean total number of steps taken per day?

Library Dplyr to work with "%>%"

```{r Library, echo=TRUE, message=FALSE}
library(dplyr)
```

The total number of steps is the sum of all steps measured by day. The mean is the sum divided by the number of intervals..

```{r Summarizing the steps grouping by day, echo=TRUE}
Act_Day<- Act %>% 
  group_by(date) %>% 
  summarise(steps = sum(steps))
```

## Including Plots

Ploting a histogram with the sum, mean and median of steps by day.

```{r ploting 1 , echo=TRUE}
hist(Act_Day$steps, 
     breaks= seq(from=0, to=25000, by=2500), 
     col="darkblue", 
     border="black", 
     xlab="Number of Steps", 
     main="Mean and Median of the total number of steps by day", 
     ylim=c(0, 20),
     xlim=c(0,25000))

abline(v=median(Act_Day$steps,
                na.rm=TRUE), 
       col="green",
       lwd=4)

abline(v=mean(Act_Day$steps, 
              na.rm=TRUE), 
       col="red",lwd=2)

legend(x="topright", #subtitle postion
      c("Median","Mean"), #name of the subtitle
      col=c("green","red"), #colors
      lty=c(1,1), #line style
      lwd=c(4,2)) #line widht
```

## What is the average daily activity pattern?

The daily activity pattern is the mean of the number of steps of all days by interval.

```{r Mean the steps grouping by interval, echo=TRUE}
Act_Interval<- Act %>% 
  group_by(interval) %>% 
  summarise(steps = mean(steps, 
                         na.rm=TRUE))
```

Ploting the Mean of steps of all days by Interval with the line plot.

```{r ploting 2, echo=TRUE}
plot(Act_Interval$interval, 
     Act_Interval$steps, 
     type="l", 
     col="darkblue", 
     xlab="Interval(Minute)", 
     ylab="Mean of steps", 
     main="The daily activity pattern")

abline(h=max(Act_Interval$steps),
       col="green",
       lwd=2)

legend(x="topright",
      "Valor máximo",
      col="green",
      lty=1,
      lwd=2)
```

## Imputing missing values

To represent the missing data and verify the number of the missing data in each column, a nice way to do it is with the package "mice" and "VIM".

```{r Counting missing data, echo=TRUE, message=FALSE}
library(mice)
library(VIM)

mice_plot <- aggr(Act_Day, 
                  col=c('navyblue','yellow'),
                  numbers=TRUE, 
                  sortVars=TRUE,
                  labels=names(Act_Day), 
                  cex.axis=.7,
                  gap=3, 
                  ylab=c("Missing data","Pattern"))
```

The result show us that is a big amount, x>5%,  of NA's in the data, we have to take care to impute the a value that help us with accuracy.

With library "mice" and the pmm method, we will impute values to missing data.

```{r Imputing values to missing data, echo=TRUE, message=FALSE, cache=TRUE}
imputed_Data <- mice(Act, 
                     m=5, 
                     maxit = 50, 
                     method = 'pmm', 
                     seed = 500)

completeData <- complete(imputed_Data,2)

completeDataSum<- completeData %>% 
  group_by(date) %>% 
  summarise(steps = sum(steps))
```

Ploting the complete data.

```{r ploting complete data, echo=TRUE, message=FALSE}
hist(completeDataSum$steps, 
     breaks= seq(from=0, to=25000, by=2500), 
     col="darkgreen", 
     border="black", 
     xlab="Number of Steps", 
     main="Mean and Median of the total number of steps by day", 
     ylim=c(0, 20),
     xlim=c(0,25000))
      

abline(v=median(completeDataSum$steps,na.rm=TRUE), 
       col="blue",
       lwd=4)

abline(v=mean(completeDataSum$steps, 
              na.rm=TRUE), 
       col="red",
       lwd=2)

legend(x="topright",
      c("Median","Mean"),
      col=c("blue","red"),
      lty=c(1,1), 
      lwd=c(4,2)) 
```

As we can see, imputing values with de pmm method showed us a similar result comparing with the first part of the assignment. So, we can assume that imputing missing values did not impacted signifcally.

## Are there differences in activity patterns between weekdays and weekends?

```{r Processing data to 2 levels, "weekdays" and "weekend" , echo=TRUE}

Act_days<- mutate(completeData, 
                  days=weekdays(completeData$date))

Act_week<- mutate(Act_days,
                  weekpart=(ifelse(Act_days$days=="sábado"|Act_days$days=="domingo","weekend","weekday")))

Act_weekdays<- Act_week %>% 
  group_by(interval,weekpart) %>% 
  summarise(steps = mean(steps))
```

Ploting the comparassion between the parts of the week, "weekdays" and "weekend", to see if have some pattern there.

```{r ploting data to 2 levels, echo=TRUE}

library(ggplot2)

ggplot(aes(x=interval,y=steps),
       data=Act_weekdays)+
       geom_line(col="blue")+
       facet_wrap(~Act_weekdays$weekpart)

```