---
title: "Statistical Inference Course Project Part 2"
author: "Luis Fernando Agottani"
date: "24/08/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For the second part, the ToothGrowth data in the R dataset package will be analyzed.

## Part 2: Basic Inferential Data Analysis Instructions

1. Load the ToothGrowth data and perform some basic exploratory data analyses

```{r Loading data, echo=TRUE}
library(datasets)

data(ToothGrowth)
```

2. Provide a basic summary of the data.

```{r Summary data, echo=TRUE}
dim(ToothGrowth)

str(ToothGrowth)

head(ToothGrowth)

summary(ToothGrowth)
```

The data ToothGrowth give to us results from testing vitamin C in Guinea Pigs to measure the tooth growth in differents dosages (0,5; 1,0; 2,0 mg) and in differents types of supplements (VC and OJ).

```{r Exploratory Analisis, echo=TRUE}
library(ggplot2)

qplot(supp,len,data=ToothGrowth, facets=~dose, main="Tooth growth of guinea pigs by supplement type and dosage (mg)",xlab="Supplement type", ylab="Tooth length") + geom_boxplot(aes(fill = supp))
```

The result of the exploratory analises from the data "ToothGrowth", show that the tooth length increases when the dosage of vitamin C bigger, and for dosages of 0,5mg and 1mg, the supplement OJ has better results when comparing to VC. Lets confirm these with hypothesis.

3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

Hypothesis to the supplement.

Lets work with Hypothesis. Using our null hypothesis when using supplement OJ and VC.

So, 
Null Ho = lenth OJ = lenght VC,
Alternative Ha = lenth OJ > lenght VC.

```{r Hypothesis lenth OJ > lenght VC, echo=TRUE}
OJ = ToothGrowth$len[ToothGrowth$supp == 'OJ']
VC = ToothGrowth$len[ToothGrowth$supp == 'VC']

t.test(OJ, VC, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```

With the graphic, we can see that for dosage equal 2mg, there is not a significant difference. Lets check.

```{r Hypothesis lenth OJ2mg > lenght VC2mg, echo=TRUE}
OJ2mg = ToothGrowth$len[ToothGrowth$supp == 'OJ' & ToothGrowth$dose == 2]
VC2mg = ToothGrowth$len[ToothGrowth$supp == 'VC' & ToothGrowth$dose == 2]

t.test(OJ2mg, VC2mg, alternative = "two.sided", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```

Hypothesis to dosage.

The null hypothesis now is that there is not difference in tooth growth between dosage and our alternative hypotesis is that when bigger the dosage greater the growth tooth effect.

```{r Hypothesis dosage, echo=TRUE}
dose05 = ToothGrowth$len[ToothGrowth$dose == '0.5']
dose10 = ToothGrowth$len[ToothGrowth$dose == '1']
dose20 = ToothGrowth$len[ToothGrowth$dose == '2']

t.test(dose10, dose05, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.95)

t.test(dose20, dose10, alternative = "greater", paired = FALSE, var.equal = FALSE, conf.level = 0.95)
```

## State your conclusions and the assumptions needed for your conclusions.

In our t.test, p-values that are lower than 5% are reject the null hypothesis. we can conclude that more dosages of vitamin C will result in bigger tooth growth (p-value<0.05) and for dosages 0,5 mg and 1,0 mg the supplement OJ has more effect than VC (p-value<0.05. For dosage 2,0 mg results shows that is not difference between supplements for tooth growth (p-value>0.05.

