---
title: "Statistical Inference Course Project - Part 1"
author: "Luis Fernando Agottani"
date: "24/08/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The first part of this project has the intention to study how the Central Limit Theorem works with means from simulations with exponential distribution.

## Part 1 - A simulation exercise.

 The goal of the first step is input the mean of each 1000 simulations with 40 exponentials, where lambda=0.2, rexp(n,lambda) and compare with de mean of the exponencial distribution, which is 1/lambda. 

```{r rexp, echo=TRUE}
n=40
simulations=1000
lambda=0.2

mns = NULL
for (i in 1 : 1000) mns = c(mns,
                            mean(rexp(n,lambda)))

```

1. Show the sample mean and compare it to the theoretical mean of the distribution.

```{r Theorical mean, echo=TRUE}
Theoricalmean<- 1/lambda

Theoricalmean

Samplemean<- mean(mns)

Samplemean
```

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r Theorical variance, echo=TRUE}
Theoricalvariance<- (1/lambda/sqrt(40))^2

Theoricalvariance

Samplevariance<- var(mns)

Samplevariance
```

As we can see, the sample mean and variance are very similar to the therorical. 

3. Show that the distribution is approximately normal.

```{r plotNORMAL, echo=TRUE}
hist(mns, 
     breaks= 25,
     prob = TRUE,
     col="grey", 
     border="white", 
     xlab="Means", 
     main="Means distribution from exponencial simulation")
  
abline(v=mean(mns), 
       col="black",
       lwd=6)

abline(v=5, 
       col="red",
       lwd=4)

legend(x="topright",
      c("Sample Mean","Theorical Mean", "Theorical Distribution"),
      col=c("black","red", "blue"),
      lty=c(1,1))

x <- seq(min(mns), max(mns), length = 100)

lines(x, dnorm(x, mean = 1/lambda, sd = (1/lambda/sqrt(n))), pch = 25, col = "blue")
```

So, the blue line represents the theorical distribuition, which is similar to the sample mean.